---
title: "traits_review_datamanip"
author: "Natasha Hardy"
date: "02/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data manipulation document

## Initial trait data

This document contains code for intial data manipulation March 2020 using large "traits_dummy.csv" df.

```{r Data Manipulation, include=F, eval=F, echo=T}

#Load data
review_traits <- read_csv("traits_dummy.csv")
#Check dataframe
#str(review_traits)
unique(review_traits$`Taxonomic`)
summary(as.factor(review_traits$`Taxonomic`))
#  Birds    Broad     Fish    Herps  Insects  Mammals Multiple    Other Plankton   Plants 
#      74      118      173       49      156       53       56      108       35      406 

#Need to rename Predictive values
#Need to rename and create some key variables for grouping factors

#Also found a typo, there are probably loads more
review_traits = review_traits %>%
  dplyr::rename(Predictive = `Forecasting/Predictive`, yield = `yeild`) %>%
  mutate(`Global Change`=if_else(review_traits$`Global Change Driver`>0, "yes", "no"), 
         `Taxonomic Group`=case_when( #Note can use case_when() or if_ele() BUT you need to write one for each value and not try to combine them as strings using c()
           review_traits$`Taxonomic` == "Mammals" ~ "Vertebrate",
           review_traits$`Taxonomic` == "Birds" ~ "Vertebrate",
           review_traits$`Taxonomic` == "Fish" ~ "Vertebrate",
           review_traits$`Taxonomic` == "Herps" ~ "Vertebrate",
           review_traits$`Taxonomic` == "Plankton" ~ "Invertebrate",
           review_traits$`Taxonomic` == "Insects" ~ "Invertebrate",
           review_traits$`Taxonomic` == "Multiple" ~ "Multiple",
           review_traits$`Taxonomic` == "Broad" ~ "Multiple",
           review_traits$`Taxonomic` == "Other" ~ "Other",
           review_traits$`Taxonomic` == "Plants" ~ "Plants")
         ) %>%
  select(`DOI`:`Taxonomic`,`Taxonomic Group`,`System`:`Global Change Driver`, `Global Change`, `# of eggs or live young`:`zoogeographical group`)
  #Add binary column for presence/absence of Global Change Driver
View(review_traits)
#Check Taxonomic group variable
#unique(review_traits$`Taxonomic Group`)
#summary(as.factor(review_traits$`Taxonomic Group`))

# DF problems: at "string 1291". This appears to be a duplicate of column #1292
sum(review_traits[,1291]) #10
colnames(review_traits[,1291]) #"locomotion\x96substrate relation"
sum(review_traits[,1292]) #4
colnames(review_traits[,1292]) #"locomotion and substrate relation"
#Need to merge these columns

# USE THIS DF
#Merging two repetitive/erroneous locomotion ~ substrate trait columns
review_traits_fix = review_traits %>%
   mutate(`locomotion and substrate relation total`=`locomotion\x96substrate relation`+`locomotion and substrate relation`) %>%
  select(-c(`locomotion\x96substrate relation`,`locomotion and substrate relation`))
#Check
#sum(review_traits_fix$`locomotion and substrate relation total`) #6

#The full data probably contains columns that may have occurred once
#First deleting any columns with zero sum column values
#review_species.0=review_species[,-which(colSums(review_species)==0)] #ncol = 2544
#Second deleting any columns with =< 1 occurrences of a trait value
#review_species.1=review_species[,-which(colSums(review_species)<2)] #ncol = 1413
```

```{r Data 1.0, include=F, eval=F, echo=T}
# Remove zero sum columns and rows from the dataframe

#Remove columns that only contain a single occurrence for a trait, as well as any zero sum rows
review_traits.1 = review_traits_fix[,-which(colSums(review_traits_fix[,11:ncol(review_traits_fix)])<2)]
#Not sure why some explanatory variable columns disappear also, so adding them back in:
review_traits.2 = cbind(review_traits_fix[,1:10], review_traits.1[,4:ncol(review_traits.1)])
#Then delete rows that sum to zero for the trait occurrences
review_traits1.0 = review_traits.2[-which(rowSums(review_traits.2[11:ncol(review_traits.2)])==0),] #1099 observations of 1415 traits

#USE THESE SUBSETS
review_sites1.0 <- review_traits1.0[,1:10]
review_species1.0 <- review_traits1.0[,11:ncol(review_traits1.0)] #1410 trait columns

```

```{r Data 2.0, include=F, eval=F, echo=T}
#DO NOT USE
#This chunk selects distinct studies by DOI from review_traits1.0, but it doesn't appear to be doing this randomly
#unique(review_traits1.0$DOI) #705

#review_traits2.0 = review_traits1.0 %>%
#  distinct(review_traits1.0$DOI, .keep_all = TRUE) %>%
#  select(DOI:`locomotion and substrate relation total`)
#Datasets
#review_sites2.0 <- review_traits2.0[,1:10]
#review_species2.0 <- review_traits2.0[,11:ncol(review_traits2.0)] #1410 trait columns
#This doesn't appear to have randomly selected studies but instead taken a lot of Abiotic studies


```

```{r Data 3.0, include=F, eval=F, echo=T}
#DO NOT USE
#Does not appear to be entirely random
#Remove any DOI that appeared twice
#review_traits2.5 = review_traits1.0 %>%
#  group_by(DOI) %>% 
#  dplyr::summarize(obs = n())

#review_traits2.7 = review_traits1.0 %>%
#  left_join(review_traits2.5) %>%
#  select(DOI, obs, Ecosystem:`Global Change`, `# of eggs or live young`:`locomotion and substrate relation total`) 

#review_traits3.0 = review_traits2.7 %>%
#  filter(as.numeric(review_traits2.7$obs) == "1")
  
#str(review_traits2.7$obs)
#summary(review_traits2.7$obs) #374
#review_traits2.7$obs <- as.factor(review_traits2.7$obs)

#summary(as.factor(review_traits3.0$Filter))

#Datasets
#review_sites3.0 <- review_traits3.0[,1:11]
#review_species3.0 <- review_traits3.0[,12:ncol(review_traits3.0)] #1410 trait columns

```

### USE THIS DATA MANIP

```{r Data 4.0, include=F, eval=F, echo=T}
#USE THESE DATA
#Randomly select studies by DOI

review_traits4.0 = review_traits1.0 %>% 
   group_by(DOI) %>%
   sample_n(1)
#str(review_traits4.0)
summary(as.factor(review_traits4.0$Filter))

#Because this randomly selects from available DOI's I have run this a few times to view a consensus on the representation of various variables (Ecosystems, Filter, etc.). Some versions appear to have randomly selected too few of a given variable's values

#Save different iterations for reuse in SIMPER and graphs!!!

#Save versions of this
#write.csv(review_traits4.0, "review_traits4.0.csv")
#Ecological Fundamental    Physical     Trophic 
#        109         438          36         144 

#write.csv(review_traits4.0, "review_traits4.0v2.csv")

# Ecological Fundamental    Physical     Trophic 
#        125         425          38         139 

review_sites4.0 <- as.data.frame(review_traits4.0[,1:10])
review_species4.0 <- review_traits4.0[,11:ncol(review_traits4.0)] #1410 trait columns


```


### Data issues dummy

There are way too many columns. Some thoughts on variable reduction:
(i) Reduction based on number of occurrences (>0 --> 2400 variables, >1 --> #1411 variables, >2 --> #144 variables)
(ii) Using fuzzy logic techniques to clean up and combine related variables
(iii) Manual cleaning



## Secondary trait classification

### Load and manipulate the data

```{r Secondary Data Load, include=F, eval=F, echo=T}
#This contains a cleaned up dataset of several hundred traits that have been "summarised" by a process they model, rather than completely disaggregated unique traits for which we had over 2000
#Load data
review_traits_2nd <- read_csv("secondary_traits_fill_dummy.csv")
#Check dataframe
glimpse(review_traits_2nd)

#summary(as.factor(review_traits_2nd$TOS))
#Note that there are 4 NA's under type of study and we may need to omit these.
#summary(as.factor(review_traits_2nd$`Taxonomic`))
#        Birds         Fish Herpetofauna      Insects      Mammals     Multiple        Other 
#         155          324           91          339          105          393          230 
#    Plankton       Plants 
#          66          614 

#Need to rename Predictive values
#Need to rename and create some key variables for grouping factors
```

```{r Secondary Data Manipulation, include=F, eval=F, echo=T}

#Also found a typo, there are probably loads more
review_traits_2use = review_traits_2nd %>%
  group_by(DOI) %>%
  sample_n(1) %>%
  dplyr::rename(Predictive = `Forecasting`) %>%
  mutate(`GlobalChangeCat`=if_else(`GlobalChange`>0, "yes", "no"), #Add binary column for presence/absence of Global Change Driver 
         `TaxonomicGroup`=case_when( #Note can use case_when() or if_ele() BUT you need to write one for each value and not try to combine them as strings using c()
           `Taxonomic` == "Mammals" ~ "Vertebrate",
           `Taxonomic` == "Birds" ~ "Vertebrate",
           `Taxonomic` == "Fish" ~ "Vertebrate",
           `Taxonomic` == "Herpetofauna" ~ "Vertebrate",
           `Taxonomic` == "Plankton" ~ "Invertebrate",
           `Taxonomic` == "Insects" ~ "Invertebrate",
           `Taxonomic` == "Multiple" ~ "Multiple",
           `Taxonomic` == "Other" ~ "Other",
           `Taxonomic` == "Plants" ~ "Plants")
         ) %>%
  select(`DOI`:`Taxonomic`,`TaxonomicGroup`,`GlobalChange`,`GlobalChangeCat`,  `Predictive`:`zinc`, -c(`web making`, `diving ability`, `palladium`)) %>%
  #Can remove: web-building and diving ability are problematic NMDS1 axis #, -c(`web making`, `diving ability`)
  #Need to leave this in
  filter(TOS != "NA")

glimpse(review_traits_2use)
#dim(review_traits_2use) #821 observations and 227 columns after removal of three problem variables

```

```{r Secondary Data Simplification}

# Remove columns and rows that sum to 1 from the dataframe

#Remove columns that only contain a single occurrence for a trait, as well as any zero sum rows
review_traits_2.1 = review_traits_2use[,which(colSums(review_traits_2use[,11:ncol(review_traits_2use)])>1)]
#str(review_traits_2.1)
#NOTE
#We go from having 220 traits observed to 182 observed more than once.
#Also note some explanatory variable columns disappear also, so adding them back in:
review_traits_2.2 = cbind(as.data.frame(review_traits_2use[,1:10]), as.data.frame(review_traits_2.1[,9:ncol(review_traits_2.1)]))

#WTF
#Error message:
#number of rows of result is not a multiple of vector length (arg 1)
#dim(review_traits_2use[,1:10])
#dim(review_traits_2.1[,9:ncol(review_traits_2.1)])
#This should fucking work TBH
#Solved needed to use as.data.frame()

#Then we need to delete rows that sum to zero for the trait occurrences
review_traits_2.3 = review_traits_2.2[-which(rowSums(review_traits_2.2[,11:ncol(review_traits_2.2)])==0),] #792 observations of 181 variables
#glimpse(review_traits_2.3)

write.csv(review_traits_2.3, "review_traits_2use.csv")

```

## Secondary traits not-filled

### Load and manipulate the data

```{r Secondary NF Data Load, include=F, eval=F, echo=T}
#This contains a cleaned up dataset of several hundred traits that have been "summarised" by a process they model, rather than completely disaggregated unique traits for which we had over 2000
#Load data
review_traits_NF <- read_csv("secondary_traits_empty_dummy.csv")
#Check dataframe
glimpse(review_traits_NF)

summary(as.factor(review_traits_NF$TOS))
#Note that there are 4 NA's under type of study and we may need to omit these.
summary(as.factor(review_traits_NF$`Taxonomic`))
#        Birds         Fish Herpetofauna      Insects      Mammals     Multiple        Other 
#         155          324           91          339          105          393          230 
#    Plankton       Plants 
#          66          614 

```

```{r Secondary Data Manipulation, include=F, eval=F, echo=T}

#Also found a typo, there are probably loads more
review_traits_NFuse = review_traits_NF %>%
  group_by(DOI) %>%
  sample_n(1) %>%
  dplyr::rename(Predictive = `Forecasting`) %>% #Need to rename Predictive values
  mutate(`GlobalChangeCat`=if_else(`GlobalChange`>0, "yes", "no"), #Add binary column for presence/absence of Global Change Driver 
         `PredictiveCat`=if_else(`Predictive`>0, "yes", "no"), ##Add binary column for presence/absence of predictive work 
         `TaxonomicGroup`=case_when( #Note can use case_when() or if_ele() BUT you need to write one for each value and not try to combine them as strings using c()
           `Taxonomic` == "Mammals" ~ "Vertebrate",
           `Taxonomic` == "Birds" ~ "Vertebrate",
           `Taxonomic` == "Fish" ~ "Vertebrate",
           `Taxonomic` == "Herpetofauna" ~ "Vertebrate",
           `Taxonomic` == "Plankton" ~ "Invertebrate",
           `Taxonomic` == "Insects" ~ "Invertebrate",
           `Taxonomic` == "Multiple" ~ "Multiple",
           `Taxonomic` == "Other" ~ "Other",
           `Taxonomic` == "Plants" ~ "Plants")
         ) %>%
  select(`DOI`:`Taxonomic`,`TaxonomicGroup`,`GlobalChange`,`GlobalChangeCat`, `Predictive`, `PredictiveCat`, `TOS`:`filter`, `age`:`zinc`) %>%
  #Can remove: , -cadmium
  #Need to leave this in
  filter(TOS != "NA")

glimpse(review_traits_NFuse)
#dim(review_traits_2use) #821 observations and 226 columns after removal of three problem variables

```

```{r Secondary Data Simplification}

# Remove columns and rows that sum to 1 from the dataframe

#Remove columns that only contain a single occurrence for a trait, as well as any zero sum rows
review_traits_NF2.1 = review_traits_NFuse[,which(colSums(review_traits_NFuse[,12:ncol(review_traits_NFuse)])>1)]
str(review_traits_NF2.1)
#cadmium removed

#Also note some explanatory variable columns disappear also, so adding them back in:
review_traits_NF2.2 = cbind(as.data.frame(review_traits_NFuse[,1:11]), as.data.frame(review_traits_NF2.1[,10:ncol(review_traits_NF2.1)]))

#Then we need to delete rows that sum to zero for the trait occurrences
review_traits_NF2.3 = review_traits_NF2.2[-which(rowSums(review_traits_NF2.2[,12:ncol(review_traits_NF2.2)])==0),] #792 observations of 181 variables
glimpse(review_traits_NF2.3)

write.csv(review_traits_NF2.3, "review_traits_NFuse.csv")

```

## Primary trait classification

### Data load and manipulation

```{r Primary data load}

#This contains a cleaned up dataset of several hundred traits that have been "summarised" by a process they model, rather than completely disaggregated unique traits for which we had over 2000
#Load data
review_traits_1ry <- read_csv("primary_traits_dummy.csv")
#Check dataframe
glimpse(review_traits_1ry)

#summary(as.factor(review_traits_2nd$TOS))
#Note that there are 4 NA's under type of study and we may need to omit these.
#summary(as.factor(review_traits_2nd$`Taxonomic`))
#        Birds         Fish Herpetofauna      Insects      Mammals     Multiple        Other 
#         155          324           91          339          105          393          230 
#    Plankton       Plants 
#          66          614 

#Need to rename Predictive values
#Need to rename and create some key variables for grouping factors
```

```{r Primary data manipulation}

#Manipulate data to relabel a couple things
review_traits_1use = review_traits_1ry %>%
  group_by(DOI) %>%
  sample_n(1) %>%
  dplyr::rename(Predictive = `Forecasting`) %>%
  mutate(`GlobalChangeCat`=if_else(`GlobalChange`>0, "yes", "no"), 
         #Add binary column for presence/absence of Global Change Driver
         `Taxonomic Group`=case_when( 
           #Note can use case_when() or if_ele() BUT you need to write one for each value and not try to combine them as strings using c()
           `Taxonomic` == "Mammals" ~ "Vertebrate",
           `Taxonomic` == "Birds" ~ "Vertebrate",
           `Taxonomic` == "Fish" ~ "Vertebrate",
           `Taxonomic` == "Herpetofauna" ~ "Vertebrate",
           `Taxonomic` == "Plankton" ~ "Invertebrate",
           `Taxonomic` == "Insects" ~ "Invertebrate",
           `Taxonomic` == "Multiple" ~ "Multiple",
           `Taxonomic` == "Other" ~ "Other",
           `Taxonomic` == "Plants" ~ "Plants")
         ) %>%
  select(`DOI`:`Taxonomic`,`Taxonomic Group`,`GlobalChange`,`GlobalChangeCat`,  `Predictive`:`survival`) %>%
  filter(TOS != "NA")

dim(review_traits_1use)

#These data do not contain zero sum columns or rows
#CREATE SUBSETS FOR ANALYSES
#review_sites_1ry <- review_traits_1use[,1:10]
#review_species_1ry <- review_traits_1use[,11:ncol(review_traits_1use)] #9 trait columns

write.csv(review_traits_1use, "review_traits_1use.csv")

```

### For analyses
Potential data manipulation

```{r Data - remove zero variance data}

#I haven't applied this to this analysis, but this is a standard procedure in multivariate stats, to remove columns of data that do not contribute to the variance between groups.
#Remove variables with zero variance
#review_data_pca <- review_species_use[,apply(review_species_use, 2, var, na.rm=TRUE) != 0]
#review_data_pca2 <- review_data_pca[-c(373,496),]

```
