---
title: "trait_review_nmds"
author: "Natasha Hardy"
date: "March 10, 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Workspace Install, echo=F, eval=F}

#getwd()
#install.packages("knitr")
#install.packages("tidyverse")
#install.packages("vegan")
#install.packages("viridis")
#install.packages("devtools")
#devtools::install_github("jakelawlor/PNWColors") 
#install.packages("mvabund")

```

```{r Workspace Load}

library(devtools)
library(knitr)
library(tidyverse)
library(vegan)
library(ggplot2)
library(viridis)
library(plyr)
library(dplyr)
library(PNWColors)
library(mvabund)
library(reshape2)

```

```{r USE THIS DATA FOR ORDS}

#Upload selected dataset

review_traits_use <- read_csv("review_traits4.0V.csv") %>%
  select(DOI:ncol(.))

review_sites_use <- as.data.frame(review_traits_use[,1:10])
review_species_use <- review_traits_use[,11:ncol(review_traits_use)] #1410 trait columns

```

```{r USE THIS DATA FOR MV GLMs}

#Upload selected dataset

#Remove columns that contain < 2 occurrences
review_traits_use2 = review_traits_use[,-which(colSums(review_traits_use[,11:ncol(review_traits_use)])<3)]
#Not sure why some explanatory variable columns disappear also, so adding them back in:
review_traits_use3 = cbind(review_traits_use[,1:10], review_traits_use2[,1:ncol(review_traits_use2)])
#Then delete rows that sum to zero for the trait occurrences
review_traits_mv = review_traits_use3[-which(rowSums(review_traits_use3[11:ncol(review_traits_use3)])==0),] #209 observations of 144 traits

review_sites_mv <- as.data.frame(review_traits_mv[,1:10])
review_species_mv <- review_traits_mv[,11:ncol(review_traits_mv)] #144 trait columns

#write.csv(review_traits_mv, "review_traits_mv.csv")
#review_traits_mv <- read_csv("review_traits4.0V.scsv")

```

## Ordination

Note that the stress decreases as we decrease the number of observations, however, I'm concerned that it becomes so low as to indicate that we do not have enough observations vs. traits (data are too wide because #traits > # obs)


```{r Ordination USE, eval=FALSE}
#Second ordination based on using no replicates for the study
traits_ord_use <- metaMDS(review_species_use,
                          distance = "jaccard",
                          trymax = 99,
                          k = 3) #increase trymax = 50

plot(traits_ord_use) #plpots species (here traits) as black dots, sites (here studies) as red crosses, not labelled

```

## Graphics pipeline
Note that for the next three chunks below, here we change the input ordination AND the grouping variables solely and obtain the same output labels for downstream plotting
1) First using traits_ord_jac1 and review_sites1.0
2) Second using traits_ord_jac2 and review_sites2.0
3) Second using traits_ord_jac3 and review_sites3.0

### Data and treatment scores
```{r Data scores for plotting - 1.0 dataset, include=F, eval=F, echo=F}
#attach(review_sites)

data.scores <- as.data.frame(scores(traits_ord_jac1))  
#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$points <- rownames(data.scores)  
# create a column of site names, from the rownames of data.scores
treatment.scores<-cbind(review_sites1.0, data.scores)
#bind treatment labels and score values

#Check
str(treatment.scores)  
```

```{r Data scores for plotting - 2.0 dataset, include=F, eval=F, echo=F}
#attach(review_sites)

data.scores <- as.data.frame(scores(traits_ord_jac2))  
#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$points <- rownames(data.scores)  
# create a column of site names, from the rownames of data.scores
treatment.scores<-cbind(review_sites2.0, data.scores)
#bind treatment labels and score values

#Check
str(treatment.scores)  
```

```{r Data scores for plotting - 3.0 dataset, include=F, eval=F, echo=F}
#attach(review_sites)

data.scores <- as.data.frame(scores(traits_ord_jac3))  
#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$points <- rownames(data.scores)  
# create a column of site names, from the rownames of data.scores
treatment.scores<-cbind(review_sites3.0, data.scores)
#bind treatment labels and score values

#Check
str(treatment.scores)  
```

```{r Data scores for plotting - 4.0 dataset, include=F, eval=F, echo=F}
#Troubleshoot
#str(data.scores4.0) #705 and 4 variables
#str(data.scores) #705 and 4 variables

#str(review_sites4.0[,1:9]) #705 and 9 variables
#str(review_sites3.0) #705 and 9 variables

data.scores <- as.data.frame(scores(traits_ord_jac4))  
#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$points <- rownames(data.scores)  
# create a column of site names, from the rownames of data.scores
treatment.scores <- cbind(review_sites4.0, data.scores)

#bind treatment labels and score values

#Check
str(treatment.scores)  

```

```{r Data scores for plotting - USE}

#Extract NMDS coordinates and associate with co-variates/grouping factors

data.scores <- as.data.frame(scores(traits_ord_use))  
#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$points <- rownames(data.scores)  
# create a column of site names, from the rownames of data.scores
treatment.scores <- cbind(review_sites_use, data.scores)

#bind treatment labels and score values

#Check
str(treatment.scores)  

```

### Convex hull calculations
For each ordination and set of grouping variables input to data scores chunk

```{r Convex hulls - Ecosystem}
# Ecosystem type
#Create convex hulls for the space occupied by each Ecosystem type
unique(treatment.scores$Ecosystem) #"Terrestrial" "Freshwater"  "Marine"      "Broad"
treatment.scores$Ecosystem <- as.factor(treatment.scores$Ecosystem)
summary(treatment.scores$Ecosystem)

#Terrestrial
grp.t <- treatment.scores[treatment.scores$Ecosystem == "Terrestrial", ][chull(treatment.scores[treatment.scores$Ecosystem == "Terrestrial", c("NMDS1", "NMDS2")]), ] 
#Freshwater
grp.f <- treatment.scores[treatment.scores$Ecosystem == "Freshwater", ][chull(treatment.scores[treatment.scores$Ecosystem == "Freshwater", c("NMDS1", "NMDS2")]), ] 
#Marine
grp.m <- treatment.scores[treatment.scores$Ecosystem == "Marine", ][chull(treatment.scores[treatment.scores$Ecosystem == "Marine", c("NMDS1", "NMDS2")]), ]
#Broad
grp.b <- treatment.scores[treatment.scores$Ecosystem == "Broad", ][chull(treatment.scores[treatment.scores$Ecosystem == "Broad", c("NMDS1", "NMDS2")]), ]
#combine the hull data
hull.data.ecos <- rbind(grp.t, grp.f, grp.m, grp.b)  
hull.data.ecos
```

```{r Convex hulls - Global Change}
#Global Change yes/no
#Create convex hulls for the space occupied by each Ecosystem type
unique(treatment.scores$`Global Change`) #no/yes
treatment.scores$`Global Change` <- as.factor(treatment.scores$`Global Change`)
summary(treatment.scores$`Global Change`)

#Yes
grp.climyes <- treatment.scores[treatment.scores$`Global Change` == "yes", ][chull(treatment.scores[treatment.scores$`Global Change` == "yes", c("NMDS1", "NMDS2")]), ] 
#No
grp.climno <- treatment.scores[treatment.scores$`Global Change` == "no", ][chull(treatment.scores[treatment.scores$`Global Change` == "no", c("NMDS1", "NMDS2")]), ] 

#combine the hull data
hull.data.clim <- rbind(grp.climyes, grp.climno)  
hull.data.clim
```

```{r Convex hulls - Taxonomic specific}
# Taxonomic groups
#Create convex hulls for the space occupied by each Taxonomic values
unique(treatment.scores$Taxonomic) 
treatment.scores$Taxonomic <- as.factor(treatment.scores$Taxonomic)
#Change some of the values
levels(treatment.scores$Taxonomic)[levels(treatment.scores$Taxonomic)=="Broad"] <- "Multiple"
levels(treatment.scores$Taxonomic)[levels(treatment.scores$Taxonomic)=="Herps"] <- "Reptiles"
#"Birds"    "Insects"  "Mammals"  "Broad"    "Multiple" "Other"    "Fish"     "Herps"    "Plants"   "Plankton"
summary(treatment.scores$Taxonomic)

#Example loop
#for (unique(treatment.scores$Taxonomic) in treatment.scores$Taxonomic){
#  print(paste("Taxon is", unique(treatment.scores$Taxonomic)))
#}

#Taxonomic group - hull loop
tax = as.character(unique(treatment.scores$Taxonomic))
for(i in 1:length(tax)) {
  temp = tax[i]
  df = treatment.scores[treatment.scores$Taxonomic == temp, ][chull(treatment.scores[treatment.scores$Taxonomic == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.taxon <- rbind(grp.Birds, grp.Insects, grp.Mammals, grp.Multiple, grp.Other, grp.Fish, grp.Reptiles, grp.Plants, grp.Plankton)  
head(hull.data.taxon)
```

```{r Convex hulls - Taxonomic Group}
# Taxonomic groups
#Create convex hulls for the space occupied by each Taxonomic values
unique(treatment.scores$`Taxonomic Group`) 

#Taxonomic group - hull loop
taxgrp = as.character(unique(treatment.scores$`Taxonomic Group`))
for(i in 1:length(taxgrp)) {
  temp = taxgrp[i]
  df = treatment.scores[treatment.scores$`Taxonomic Group` == temp, ][chull(treatment.scores[treatment.scores$`Taxonomic Group` == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.taxgrp <- rbind(grp.Vertebrate, grp.Invertebrate, grp.Multiple, grp.Other, grp.Plants)  
head(hull.data.taxgrp)
```

```{r Convex hulls - Type of Study}
# Type of study
unique(treatment.scores$TOS)
#TModel --> Theory
treatment.scores$TOS <- as.factor(treatment.scores$TOS)
levels(treatment.scores$TOS)[levels(treatment.scores$TOS)=="TModel"] <- "Theory"
#Create convex hulls for the space occupied by each Type of study
#Hull loop
tos = as.character(unique(treatment.scores$TOS))
for(i in 1:length(tos)) {
  temp = tos[i]
  df = treatment.scores[treatment.scores$TOS == temp, ][chull(treatment.scores[treatment.scores$TOS == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.tos <- rbind(grp.Observational, grp.Experiment, grp.Metanalysis, grp.Theory, grp.Review)  
```

```{r Convex hulls - Filter}
# Level of filtering

#Make following value changes
#"Fundamental" --> Abiotic
#"Physical" --> Dispersal 
#"Ecological"  --> Biotic
#"Trophic" = "Trophic"
treatment.scores$Filter <- as.factor(treatment.scores$Filter)
levels(treatment.scores$Filter)[levels(treatment.scores$Filter)=="Fundamental"] <- "Abiotic"
levels(treatment.scores$Filter)[levels(treatment.scores$Filter)=="Physical"] <- "Dispersal"
levels(treatment.scores$Filter)[levels(treatment.scores$Filter)=="Ecological"] <- "Biotic"

#Create convex hulls for the space occupied by each Type of study
#Hull loop
filt = as.character(unique(treatment.scores$Filter))
for(i in 1:length(filt)) {
  temp = filt[i]
  df = treatment.scores[treatment.scores$Filter == temp, ][chull(treatment.scores[treatment.scores$Filter == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}


#combine the hull data
hull.data.filt <- rbind(grp.Abiotic, grp.Biotic, grp.Trophic, grp.Dispersal)  
unique(hull.data.filt$Filter)
summary(treatment.scores$Filter)

```


## nMDS Plots

### nMDS plot for the assessmenet of global change (yes/no) and the drivers.

```{r nMDS Global Change Driver, echo=TRUE}

#Change one driver label
treatment.scores$`Global Change Driver` <- as.factor(treatment.scores$`Global Change Driver`)
levels(treatment.scores$`Global Change Driver`)[levels(treatment.scores$`Global Change Driver`)==0] <- "Not Assessed"
hull.data.clim$`Global Change` <- as.factor(hull.data.clim$`Global Change`)
levels(hull.data.clim$`Global Change`)[levels(hull.data.clim$`Global Change`)=="no"] <- "No"
levels(hull.data.clim$`Global Change`)[levels(hull.data.clim$`Global Change`)=="yes"] <- "Yes"
#Make palette for points?
#Make palette for polygons?


trait_nMDS_clim <- ggplot() + 
  geom_polygon(data=hull.data.clim, aes(x=NMDS1,y=NMDS2, fill=`Global Change`, group=`Global Change`), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores, aes(x=NMDS1,y=NMDS2, colour = `Global Change Driver`), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_fill_viridis(discrete = TRUE, begin = 1, end = 0.4, name = "Global Change Assessed") +
  scale_colour_viridis(discrete = TRUE, begin = 1, end = 0.2)
#Need to add labels
trait_nMDS_clim

#Note, could use "System" variable for labelling

ggsave("trait_nMDS_clim.jpeg", plot = trait_nMDS_clim, width = 8, height = 8, dpi = 300)


```

### nMDS plot for the ecosystem type studied

```{r nMDS Ecosystem, echo=TRUE}

pnwpal <- pnw_palette(4, name = "Starfish", type = "continuous")

#levels(hull.data.ecos$Ecosystem) <- fct_rev(factor(hull.data.ecos$Ecosystem))
#hull.data.ecos$Ecosystem <- factor(hull.data.ecos$Ecosystem, levels = c("Terrestrial", "Freshwater", "Marine", "Broad"))

#treatment.scores$Ecosystem <- as.factor(treatment.scores$Ecosystem)
#treatment.scores$Ecosystem <- factor(treatment.scores$Ecosystem, levels = c("Terrestrial", "Freshwater", "Marine", "Broad"))
#treatment.scores$Ecosystem <- as.character(treatment.scores$Ecosystem)

trait_nMDS_ecos <- ggplot() + 
  geom_polygon(data=hull.data.ecos,aes(x=NMDS1,y=NMDS2, fill=fct_rev(`Ecosystem`), group=fct_rev(`Ecosystem`)), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_rev(`Ecosystem`)), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(si'ze=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guide_legend(title="Ecosystem") +
  scale_colour_manual(values=pnwpal, name="Ecosystem") +
  scale_fill_manual(values=pnwpal, name="Ecosystem")
  #scale_colour_viridis(discrete = TRUE, begin = 0.2, end = 1) +
  #scale_fill_viridis(discrete = TRUE, begin = 0.2, end = 1)
#Need to add labels

trait_nMDS_ecos

#Note, could use "System" variable for labelling

ggsave("trait_nMDS_ecos.jpeg", plot = trait_nMDS_ecos, width = 8, height = 8, dpi = 300)


```

### nMDS plot for the specific taxonomic group studied

```{r nMDS Taxonomic, echo=TRUE}

trait_nMDS_taxon <- ggplot() + 
  geom_polygon(data=hull.data.taxon,aes(x=NMDS1,y=NMDS2, fill= `Taxonomic`, group= `Taxonomic`),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = `Taxonomic`), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_colour_viridis(discrete = TRUE, name="Taxonomic") +
  scale_fill_viridis(discrete = TRUE, name="Taxonomic")
#Need to add labels

trait_nMDS_taxon

#Note, could use "System" variable for labelling

ggsave("trait_nMDS_taxonsp.jpeg", plot = trait_nMDS_taxon, width = 8, height = 8, dpi = 300)

```

### nMDS plot for the broad taxonomic group studied

```{r nMDS Taxonomic Group, echo=TRUE}

trait_nMDS_taxon <- ggplot() + 
  geom_polygon(data=hull.data.taxgrp,aes(x=NMDS1,y=NMDS2, fill= fct_rev(`Taxonomic Group`), group= fct_rev(`Taxonomic Group`)),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_rev(`Taxonomic Group`)), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_colour_viridis(discrete = TRUE, name="Taxonomic Group") +
  scale_fill_viridis(discrete = TRUE, name="Taxonomic Group")
#Need to add labels

trait_nMDS_taxon

#Note, could use "System" variable for labelling

ggsave("trait_nMDS_taxongrp.jpeg", plot = trait_nMDS_taxon, width = 8, height = 8, dpi = 300)

```

### nMDS plot for the type of study

```{r nMDS Type of study, echo=TRUE}

#pnwpal2 <- pnw_palette(5, name = "Anemone", type = "continuous")

trait_nMDS_tos <- ggplot() + 
  geom_polygon(data=hull.data.tos,aes(x=NMDS1,y=NMDS2,fill=TOS,group=TOS),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = TOS), size=2) + # add the point markers
  coord_equal() +
  #theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18), 
        panel.background = element_rect(fill = "lightgrey"), #element_blank()
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_colour_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE)
  #scale_colour_manual(values=pnwpal2) +
  #scale_fill_manual(values=pnwpal2)
#Need to add labels

trait_nMDS_tos

#Note, could use "System" variable for labelling

ggsave("trait_nMDS_tos2.jpeg", plot = trait_nMDS_tos, width = 8, height = 8, dpi = 300)

```

### nMDS plot for the level of ecological filtering for which traits were applied

```{r nMDS Filter, echo=TRUE}

pnwpal <- pnw_palette(4, name = "Starfish", type = "continuous")
#Change order of variables for polygons
levels(hull.data.filt$Filter)
#hull.data.filt$Filter <- factor(hull.data.filt$Filter, levels = c("Abiotic", "Biotic", "Trophic", "Dispersal"))
#hull.data.filt$Filter <- factor(hull.data.filt$Filter, levels = c("Dispersal", "Trophic", "Biotic", "Abiotic"))
#Change order of variables for points
levels(treatment.scores$Filter)
#treatment.scores$Filter <- factor(treatment.scores$Filter, levels = c("Abiotic", "Biotic", "Trophic", "Dispersal"))
#treatment.scores$Filter <- factor(treatment.scores$Filter, levels = c("Dispersal", "Trophic", "Biotic", "Abiotic"))

trait_nMDS_filt <- ggplot() + 
  geom_polygon(data=hull.data.filt,aes(x=NMDS1,y=NMDS2,fill=Filter,group=Filter),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = Filter), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18), 
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_colour_manual(values=pnwpal, name="Filter") +
  scale_fill_manual(values=pnwpal, name="Filter")
  #scale_colour_viridis(discrete = TRUE, begin = 0.2, end = 1) +
  #scale_fill_viridis(discrete = TRUE, begin = 0.2, end = 1)
#Need to add labels

trait_nMDS_filt

#Note, could use "System" variable for labelling

ggsave("trait_nMDS_filt.jpeg", plot = trait_nMDS_filt, width = 8, height = 8, dpi = 300)

```

## Simper Analyses

Notes from rdocumentation: https://www.rdocumentation.org/packages/vegan/versions/2.4-2/topics/simper

The simper functions performs pairwise comparisons of groups of sampling units and finds the average contributions of each species to the average overall Bray-Curtis dissimilarity.

The function displays most important species for each pair of groups. These species contribute at least to 70 % of the differences between groups. The function returns much more extensive results which can be accessed directly from the result object (see section Value). Function summary transforms the result to a list of data frames. With argument ordered = TRUE the data frames also include the cumulative contributions and are ordered by species contribution.

The results of simper can be very difficult to interpret. The method very badly confounds the mean between group differences and within group variation, and seems to single out variable species instead of distinctive species (Warton et al. 2012). Even if you make groups that are copies of each other, the method will single out species with high contribution, but these are not contributions to non-existing between-group differences but to within-group variation in species abundance.

```{r SIMPER Ecosystem}

simper_ecos <- simper(review_species_use, review_sites_use$Ecosystem, permutations = 1000)

summary_ecos <- summary(simper_ecos)

write.csv(simper_ecos$Freshwater_Terrestrial, "simper_ecos_Freshwater_Terrestrial.csv")
write.csv(simper_ecos$Freshwater_Marine, "simper_ecos_Freshwater_Marine.csv")
write.csv(simper_ecos$Freshwater_Broad, "simper_ecos_Freshwater_Broad.csv")
write.csv(simper_ecos$Terrestrial_Marine, "simper_ecos_Terrestrial_Marine.csv")
write.csv(simper_ecos$Terrestrial_Broad, "simper_ecos_Terrestrial_Broad.csv")
write.csv(simper_ecos$Marine_Broad, "simper_ecos_Marine_Broad.csv")

```

```{r SIMPER Filter}

simper_filter <- simper(review_species_use, review_sites_use$Filter, permutations = 1000)

summary_filter <- summary(simper_filter)

write.csv(simper_filter$Ecological_Fundamental, "simper_filter_Ecological_Fundamental.csv")
write.csv(simper_filter$Ecological_Physical, "simper_filter_Ecological_Physical.csv")
write.csv(simper_filter$Ecological_Trophic, "simper_filter_Ecological_Trophic.csv")
write.csv(simper_filter$Fundamental_Physical, "simper_filter_Fundamental_Physical.csv")
write.csv(simper_filter$Fundamental_Trophic, "simper_filter_Fundamental_Trophic.csv")
write.csv(simper_filter$Physical_Trophic, "simper_filter_Physical_Trophic.csv")

```

```{r SIMPER Global Change}

simper_gc <- simper(review_species_use, review_sites_use$`Global Change`, permutations = 1000)

summary_gc <- summary(simper_gc)

write.csv(simper_gc$yes_no, "simper_gc_yes_no.csv")

```

```{r SIMPER Type of Study}

simper_tos <- simper(review_species_use, review_sites_use$TOS, permutations = 1000)

summary_tos <- summary(simper_tos)

write.csv(simper_tos$Observational_Experiment, "simper_tos_Observational_Experiment.csv")
write.csv(simper_tos$Observational_Review, "simper_tos_Observational_Review.csv")
write.csv(simper_tos$Observational_Metanalysis, "simper_tos_Observational_Metanalysis.csv")
write.csv(simper_tos$Observational_TModel, "simper_tos_Observational_TModel.csv")
write.csv(simper_tos$Experiment_Review, "simper_tos_Experiment_Review.csv")
write.csv(simper_tos$Experiment_Metanalysis, "simper_tos_Experiment_Metanalysis.csv")
write.csv(simper_tos$Experiment_TModel, "simper_tos_Experiment_TModel.csv")
write.csv(simper_tos$Review_Metanalysis, "simper_tos_Review_Metanalysis.csv")
write.csv(simper_tos$Review_TModel, "simper_tos_Review_TModel.csv")
write.csv(simper_tos$Metanalysis_TModel, "simper_tos_Metanalysis_TModel.csv")

```


```{r SIMPER Predictive}

simper_predict <- simper(review_species_use, review_sites_use$Predictive, permutations = 1000)

#summary_tos <- summary(simper_tos)

write.csv(simper_predict$`0_1`, "simper_PREDICT.csv")

```

## ManyGLMs to assess differences between groups

```{r ManyGLM Setup}

#Make mvabund ojbect
review_species_mv <- mvabund(review_species_mv)

#Issue with full data
#Error in make.names(vnames, unique = TRUE) : invalid multibyte string 1211
#review_species_mv <- mvabund(review_species_use[,-1211])
#problematic column with only 2 occurrence of this trait, deleted it.

```

```{r ManyGLM A Global Change}
#Total run time of chunk ~ 1h
##Model build
mv_gc <-manyglm(review_species_mv ~ review_traits_mv$`Global Change`, data= review_traits_mv,
                    family=binomial("cloglog")) #Build model for difference in publication trait assemblages based on assessment of global change factors
saveRDS(mv_gc, "mv_gc.rds") #Save model

##Model check
plot(mv_gc) #check residual vs. fitted values
qqnorm(residuals(mv_gc)[which(residuals(mv_gc)<10000)]); abline(c(0,1,col="red")) #check normality and fit
#model fit looking good!

##Model output significance test
mv_gc.an <- anova.manyglm(mv_gc)
saveRDS(mv_gc.an, "mv_gc_anova.rds") #Save model
write.csv(mv_gc.an$table, "mv_gc_anova_table.csv") #save ANOVA table output
#Test was significant
#Note we don't want to use the summary call or Wald stat for presence/absence data

##Individual adjusted p-values for species/traits
mv_gc.an.uni <- anova.manyglm(mv_gc,p.uni="adjusted") #Get the univariate p-values for univariate tests
#View(mv_gc.an.uni$uni.test)
saveRDS(mv_gc.an.uni, "mv_gc_univs.rds") #Save model

#Get the direction of effect fof each species with the main effect
gc_coef <- coef(mv_gc) #save to R object

```

```{r ManyGLM A Coefs Global Change}
##Note could write loop and workflow for the calculation of deviance explained

##Top species
mv_gc.s <- sort(mv_gc.an$uni.test[2,],decreasing=T,index.return=T)[1:25] #sort and select top species/traits
mv_gc.s$ix[1:25] #the column numbers of the top most impacted spp/traits

##Deviance explained
#Need > 50% deviance explainaed --> result = 25 traits explain > 50% deviance
sum(mv_gc.an$uni.test[2,mv_gc.s$ix[1:25]])*100/sum(mv_gc.an$uni.test[2,]) #25 species explained = 51.93049% Deviance

##Save as dataframe
gc_top <- as.data.frame(dimnames(review_species_mv)[[2]][mv_gc.s$ix[1:25]]) #df with the names of the top 20 traits

str(gc_top)
names(gc_top)[names(gc_top)== "dimnames(review_species_mv)[[2]][mv_gc.s$ix[1:25]]"] <- "traits" #Rename column vector of traits
#How much deviance explained?
write.csv(gc_top, "mv_gc_top.csv")

###Combine traits with their coeffs and p-values
#Creat df to combine coef values, ideally also p-values from univ anovas and the top 20
gc_coef_l <- as.data.frame(t(gc_coef)) #first transpose coef_filter
gc_coef_l$traits <- rownames(gc_coef_l) #convert rownames to a column

gc_top_coeffs <- join(gc_top, gc_coef_l) #plyr left joint to obtain coeffs of top 20 traits

colnames(gc_top_coeffs)

gc_top_coeffs = gc_top_coeffs %>%
  dplyr::rename(coef_intercept = "(Intercept)", coef_gc_yes = "review_traits_mv$`Global Change`yes") #Clean up column names

#need to join with test statistic values
gc_an_test <- as.data.frame(t(mv_gc.an.uni$uni.test)) #first transpose coef_filter
gc_an_test$traits <- rownames(gc_an_test) #convert rownames to a column
gc_top_coeffs <- join(gc_top_coeffs, gc_an_test) #join with test statistic (deviance explained)
gc_top_coeffs = gc_top_coeffs %>%
  select(-"(Intercept)") %>%
  dplyr::rename(deviance_explained = "review_traits_mv$`Global Change`") #Clean up column names

#need to join with p-values
gc_an_pvalue <- as.data.frame(t(mv_gc.an.uni$uni.p)) #first transpose coef_filter
gc_an_pvalue$traits <- rownames(gc_an_pvalue) #convert rownames to a column
gc_top_coeffs <- join(gc_top_coeffs, gc_an_pvalue) #join with test statistic (deviance explained)
gc_top_coeffs = gc_top_coeffs %>%
  select(-"(Intercept)") %>%
  dplyr::rename(p_value = "review_traits_mv$`Global Change`") #Clean up column names

write.csv(gc_top_coeffs, "gc_top_coefs.csv")

```

```{r ManyGLM B Predictive}

#Change Predictive values:
str(review_traits_mv)
levels(review_traits_mv$Predictive)[levels(review_traits_mv$Predictive)==0] <- "No"
levels(review_traits_mv$Predictive)[levels(review_traits_mv$Predictive)==1] <- "Yes"

##Model build
mv_predict <-manyglm(review_species_mv ~ review_traits_mv$Predictive, data= review_traits_mv,
                    family=binomial("cloglog")) #Build model for difference in publication trait assemblages based on assessment of global change factors
saveRDS(mv_predict, "mv_predict.rds") #Save model

##Model check
par(mfrow=c(1,2))
plot(mv_predict) #check residual vs. fitted values
qqnorm(residuals(mv_predict)[which(residuals(mv_predict)<10000)]) 
#abline(c(0,1,col="red")) #check normality and fit
#model fit looking good!

##Model output significance test
mv_predict.an <- anova.manyglm(mv_predict) #Not significant, but likely because predictive sample size too small
saveRDS(mv_predict.an, "mv_predict_anova.rds") #Save model
write.csv(mv_predict.an$table, "mv_predict_anova_table.csv") #save ANOVA table output
#Test was significant
#Note we don't want to use the summary call or Wald stat for presence/absence data

##Individual adjusted p-values for species/traits
mv_predict.an.uni <- anova.manyglm(mv_predict,p.uni="adjusted") #Get the univariate p-values for univariate tests
#View(mv_predict.an.uni$uni.test)
saveRDS(mv_predict.an.uni, "mv_predict_univs.rds") #Save model

#Get the direction of effect fof each species with the main effect
predict_coef <- coef(mv_predict) #save to R object


```

```{r ManyGLM B Coeffs Predictive}

#The output from mv model was not informative because there are too few studies that were predictive and retained in these analyses.
#Therefore think it more informative to extract just the traits that were found in predictive studies

review_traits_predict = review_traits_use %>%
  filter(Predictive != 0)
  
review_traits_predict2 = review_traits_predict[,11:ncol(review_traits_predict)]
review_traits_predict3 = review_traits_predict2[,-which(colSums(review_traits_predict2)==0)]

predict_trait_sums <- as.data.frame(colSums(review_traits_predict3))
#review_traits_predict3$traits <- rownames(review_traits_predict3)
write.csv(predict_trait_sums, "predict_traits.csv")

review_traits_predictB = review_traits_use %>%
  select(Predictive, `# of eggs or live young`:ncol(review_traits_use)) %>%
  mutate(Predictive=if_else(review_traits_use$Predictive>0, "Yes", "No")) %>%
  melt(id.vars="Predictive") %>%
  filter(value > 0) 

review_traits_predicttable = review_traits_predictB %>%
  reshape2::dcast(variable + value ~ Predictive)

write.csv(review_traits_predicttable, "predict_traits_sums.csv")

```

```{r ManyGLM Filter}
#Total run time of chunk ~ 1h

mv_filter <-manyglm(review_species_mv ~ review_traits_mv$Filter, data= review_traits_mv,
                    family=binomial("cloglog"))

saveRDS(mv_filter, "mv_filter.rds") #Save model
rds <- readRDS("mv_filter.rds")

#plot(mv_filter)

mv_filter.an <- anova.manyglm(mv_filter)

View(mv_filter.an$table)
write.csv(mv_filter.an$table, "mv_filter_anova_table.csv")
saveRDS(mv_filter.an, "mv_filter_anova.rds")

#Significant p <0.001 ***

#Output
#Analysis of Deviance Table

#Model: review_species_mv ~ review_traits_mv$Filter

#Multivariate test:
#                        Res.Df Df.diff   Dev Pr(>Dev)   
#(Intercept)                208                          
#review_traits_mv$Filter    205       3 300.9    0.005 **
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
#Arguments:
# Test statistics calculated assuming uncorrelated response (for faster computation) 
# P-value calculated using 999 iterations via PIT-trap resampling.

#Note we don't want to use the summary call or Wald stat for presence/absence data

#Get the univariate p-values for univariate tests
mv_filter.an.uni <- anova.manyglm(mv_filter,p.uni="adjusted")
View(mv_filter.an.uni$uni.p)

#Significant response variables?
saveRDS(mv_filter.an.uni, "mv_filter_univs.rds")

#Get the direction of effect for each species with the main effect
coef_filter <- coef(mv_filter)
View(coef_filter_l)
write.csv(coef_filter, "mv_filter_coefs.csv")

#Creat df to combine coef values, ideally also p-values from univ anovas and the top 20
coef_filt_l <- as.data.frame(t(coef_filter)) #first transpose coef_filter
coef_filt_l$traits <- rownames(coef_filt_l) #convert rownames to a column

str(coef_filt_l)

rownames(coef_filter)

##Top species
mv_filter.s <- sort(mv_filter.an$uni.test[2,],decreasing=T,index.return=T)[1:20]
#an$uni.test stores univariate test stats, 3rd row has change in deviance due to time
mv_filter.s$ix[1:20] # the column numbers of the "top 10" most impacted spp # so column number 25 had the strongest effect, most disappeared between the two time periods

filter_top20 <- dimnames(review_species_mv)[[2]][mv_filter.s$ix[1:20]] #the names of the top 10 spp
View(filter_top20)
write.csv(filter_top20, "mv_filter_top20.csv")
#Get coeffs for these
# [1] "diaspore.type"                           "locomotion.and.substrate.relation.total"
# [3] "vegetative.reproduction"                 "spinescence"                            
# [5] "litter.size"                             "seed.dry.mass"                          
# [7] "life.history.strategy"                   "width"                                  
# [9] "body.form"                               "swim.factor"                            
# [11] "biomass"                                 "insect.adult.longevity"                 
# [13] "larval.longevity"                        "adult.mass"                             
# [15] "defence.mechanism"                       "nest.form"                              
# [17] "development"                             "worker.body.size"                       
# [19] "method.of.locomotion"                    "length.of.the.femur"         

```


```{r ManyGLM Type of Study}
#Total run time of chunk ~ 1h

mv_tos <-manyglm(review_species_mv ~ review_traits_mv$TOS, data= review_traits_mv, 
                 family=binomial("cloglog"))

saveRDS(mv_tos, "mv_tos.rds") #Save model

#plot(mv_tos)

mv_tos.an <- anova.manyglm(mv_tos)
saveRDS(mv_tos.an, "mv_tos_anova.rds") #Save model

#Analysis of Deviance Table
#Model: review_species_mv ~ review_traits_mv$TOS
#Multivariate test:
#                     Res.Df Df.diff   Dev Pr(>Dev)  
#(Intercept)             208                         
#review_traits_mv$TOS    204       4 181.4    0.027 *
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
#Arguments:
# Test statistics calculated assuming uncorrelated response (for faster computation) 
# P-value calculated using 999 iterations via PIT-trap resampling.

#Uncomment for outputs
#Get model summary
summv_tos <- summary(mv_tos)
saveRDS(summv_tos, "mv_tos_summary.rds") #Save model

#Test statistics:
#                                  wald value Pr(>wald)   
#(Intercept)                           11.525    0.0020 **
#review_traits_mv$TOSMetanalysis        2.038    0.0689 . 
#review_traits_mv$TOSObservational      3.337    0.0110 * 
#review_traits_mv$TOSReview             1.999    0.0240 * 
#review_traits_mv$TOSTModel             1.055    0.2647   
#Test statistic:   6.34, p-value: 0.045 
#Arguments:
# Test statistics calculated assuming response assumed to be uncorrelated 
# P-value calculated using 1000 resampling iterations via pit.trap resampling (to account for correlation in #testing).

#Get the univariate p-values for univariate tests
mv_tos.an.uni <- anova.manyglm(mv_tos,p.uni="adjusted")
#Significant?
saveRDS(mv_tos.an.uni, "mv_tos_univs.rds") #Save model

```

```{r ManyGLM Ecosystem}
#Total run time of chunk ~ 1h

mv_ecos <-manyglm(review_species_mv ~ review_traits_mv$Ecosystem, data= review_traits_mv, 
                 family=binomial("cloglog"))

saveRDS(mv_ecos, "mv_ecos.rds") #Save model

#plot(mv_ecos)

mv_ecos.an <- anova.manyglm(mv_ecos)
saveRDS(mv_ecos.an, "mv_ecos_anova.rds") #Save model

#Uncomment for outputs
#Get model summary
summv_ecos <- summary(mv_ecos)
saveRDS(summv_ecos, "mv_ecos_summary.rds") #Save model

#Get the univariate p-values for univariate tests
mv_ecos.an.uni <- anova.manyglm(mv_ecos,p.uni="adjusted")
#Significant?
saveRDS(mv_ecos.an.uni, "mv_ecos_univs.rds") #Save model

```

