---
title: "traits_review_nmds_1"
author: "Natasha Hardy"
date: "03/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Workspace Install, include=FALSE}

getwd()
#install.packages("xlsx")
#library(devtools)
#install_github("vqv/ggbiplot")
#install.packages("goeveg")

```

```{r Workspace Load, include=FALSE}

#Data manip
library(tidyverse)
library(plyr)

#Load after
library(dplyr)
library(openxlsx)

#Stats / ordination
library(permute)
library(lattice)
library(vegan)
library(mvabund)

#Graphs
library(ggplot2)
library(scales)
library(grid)
library(gridExtra)
library(viridis)
library(PNWColors)
library(cowplot)
library(forcats)
library(goeveg)

```

## UPLOAD CLEAN DATA FROM REPO

UPLOAD CLEANED DATA HERE & USE THIS DATA


```{r Primary Dataset}

#Upload selected dataset

review_traits_1use <- read_csv("review_traits_1use.csv")
review_traits_1use = review_traits_1use %>%
  select(DOI:ncol(.))
review_sites_1use <- as.data.frame(review_traits_1use[,1:10])
review_species_1use <- review_traits_1use[,11:ncol(review_traits_1use)] #1410 trait columns

head(review_traits_1use)
```

## Ordinations

Using Jaccard dissimilarity for binary data.

Note that the stress decreases as we decrease the number of observations, however, I'm concerned that it becomes so low as to indicate that we do not have enough observations vs. traits (data are too wide because #traits > # obs)

### Ordination 1ry

```{r Ordination USE, echo=F, eval=F}
#Don't run, run time = 3h

#Ordination on primary trait classification 
traits_ord_1use <- metaMDS(review_species_1use,
                          distance = "jaccard",
                          trymax = 1000,
                          k = 3) #increase trymax = 50

#Based on ordination goodness fit/stress checks, we could use k = 2.
#Don't want to redo this right now as it will take ~2h to rerun.

plot(traits_ord_1use) #plots species (here traits) as black dots, sites (here studies) as red crosses, not labelled

#Plot sites (row numbers) and species (traits) as labels
ordiplot(traits_ord_1use, type = "none") #creates blank plot
orditorp(traits_ord_1use, display = "species", col = "black", air=0.01) #labels traits
orditorp(traits_ord_1use, display = "sites", col = "red", air=0.01) #labels traits


saveRDS(traits_ord_1use, "traits_ord_1use.rds")

```

### UPLOAD ORD FILE

--> You can start directly here for further analysis pipeline

```{r UPLOAD ORDINATION RDS FILE}

#primary trait classification filled in
traits_ord_1use <- readRDS("traits_ord_1use.rds")

```

## Checking ordination/nMDS output

This is a helpful tutorial on checking nMDS goodness of fit, model convergence, stress and dimensionality: https://ourcodingclub.github.io/tutorials/ordination/ 
Notes:
For the initial dummy dataset - stress is extremely low, likely there are insufficient data (n = 727 studies/rows) for the number of traits (n = 1411) included in this analysis.
For the primary traits dataset - the inflection point on the stress vs dimensions plot is between 2-3, so I am selecting k= 3 

```{r Ordination - checks}
#Run this for each ordination
#Check fit
gof <- goodness(traits_ord)

#Can visualise goodness of fit
stressplot(traits_ord, pch = 19, cex=0.75, l.col = "tomato", p.col = "skyblue")

#Now, we will be computing the stress for the dimensions 1, 2, 3, 4, 5, …. and explain how many dimensions are required to obtain a good fit.
#REMEMBER TO ADJUST THE INITIAL SP MATRIX
stress_vec <- numeric(10)
for(i in seq(10)){
  stress_vec[i] <- metaMDS(review_species_1use,
                          distance = "jaccard",
                          k=i)$stress
}

#Plot the stress vs. number of dimensions needed
plot(seq(10),stress_vec, type = 'o', ylab = "Stress", xlab = "Number of dimensions",
     col="tomato", pch=19)




#Plot goodness of fit of points
#plot(traits_ord, display = "sites", type = "n")
#points(traits_ord)

```

```{r nMDS stress check, eval=FALSE}
#Don't run, too time intensive.
#We know from output above that the model is underdispersed and we have fit too few rows vs. variables.

#dimcheckMDS(review_species_use,
#            distance = "jaccard",
#            trymax = 1000,
#            k = 3)

```


## GO TO GRAPHICS PIPELINE

For calculating data and treatment scores, and convex hulls around datapoints.

### Extract NMDS coordinates and associate with co-variates/grouping factors

```{r Data scores for plotting - USE}

#Extract NMDS coordinates and associate with co-variates/grouping factors

#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores <- as.data.frame(scores(traits_ord_1use))  

#create a column of site names, from the rownames of data.scores
data.scores$points <- rownames(data.scores)

#bind treatment labels and score values
treatment.scores <- cbind(review_sites_1use, data.scores)

#Check
str(treatment.scores)  
#Awesome

```

### Convex hull calculations
For each ordination and set of grouping variables input to data scores chunk

```{r Convex hulls - Global Change}
#Global Change yes/no
#Create convex hulls for the space occupied by Global Change yes/no
unique(treatment.scores$`GlobalChangeCat`) #no/yes
treatment.scores$`GlobalChangeCat` <- as.factor(treatment.scores$`GlobalChangeCat`)
summary(treatment.scores$`GlobalChangeCat`)
levels(treatment.scores$`GlobalChangeCat`)[levels(treatment.scores$`GlobalChangeCat`)=="no"] <- "No"
levels(treatment.scores$`GlobalChangeCat`)[levels(treatment.scores$`GlobalChangeCat`)=="yes"] <- "Yes"

#Yes
grp.climyes <- treatment.scores[treatment.scores$`GlobalChangeCat` == "Yes", ][chull(treatment.scores[treatment.scores$`GlobalChangeCat` == "Yes", c("NMDS1", "NMDS2")]), ] 
#No
grp.climno <- treatment.scores[treatment.scores$`GlobalChangeCat` == "No", ][chull(treatment.scores[treatment.scores$`GlobalChangeCat` == "No", c("NMDS1", "NMDS2")]), ] 

#combine the hull data
hull.data.clim <- rbind(grp.climyes, grp.climno)  
str(hull.data.clim)
```

```{r Convex hulls - Global Change Driver}
#Global Change Drivers
#Create convex hulls for the space occupied by each Global Change Driver
unique(treatment.scores$`GlobalChange`)

#Change driver labels
treatment.scores$`GlobalChange` <- as.factor(treatment.scores$`GlobalChange`)
summary(treatment.scores$`GlobalChange`)

levels(treatment.scores$`GlobalChange`)[levels(treatment.scores$`GlobalChange`)==0] <- "Not assessed"
levels(treatment.scores$`GlobalChange`)[levels(treatment.scores$`GlobalChange`)=="Global Change Broad"] <- "Multiple"
levels(treatment.scores$`GlobalChange`)[levels(treatment.scores$`GlobalChange`)=="Global Change Multiple"] <- "Multiple"
levels(treatment.scores$`GlobalChange`)[levels(treatment.scores$`GlobalChange`)=="Habitat Degredation"] <- "Habitat loss"
levels(treatment.scores$`GlobalChange`)[levels(treatment.scores$`GlobalChange`)=="Climate Change"] <- "Climate change" #for some reason we need to run this row, probably because it's a compound name

#Global Change Drivers - hull loop
gcd = as.character(unique(treatment.scores$`GlobalChange`))
for(i in 1:length(gcd)) {
  temp = gcd[i]
  df = treatment.scores[treatment.scores$`GlobalChange` == temp, ][chull(treatment.scores[treatment.scores$`GlobalChange` == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.gcd <- rbind(`grp.Not assessed`,`grp.Climate change`, grp.Exploitation, grp.Multiple, `grp.Habitat loss`, grp.Invasion)  

str(hull.data.gcd)

```

```{r Convex hulls - Filter}
# Level of filtering
str(treatment.scores)

treatment.scores$filter <- as.factor(treatment.scores$filter)
unique(treatment.scores$filter) #values have been changed in original df

#Create convex hulls for the space occupied by each Type of study
#Hull loop
filt = as.character(unique(treatment.scores$filter))
for(i in 1:length(filt)) {
  temp = filt[i]
  df = treatment.scores[treatment.scores$filter == temp, ][chull(treatment.scores[treatment.scores$filter == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.filt <- rbind(grp.Abiotic, grp.Biotic, grp.Trophic, grp.Dispersal)  
str(hull.data.filt)

#Check data distribution
#unique(hull.data.filt$Filter)
#summary(treatment.scores$Filter)

```

```{r Convex hulls - Predictive}
#Predictive/Forecasting studies
#Create convex hulls for the space occupied by Predictive yes/no
unique(treatment.scores$Predictive) #no/yes

#Modification of any values
treatment.scores$Predictive <- as.factor(treatment.scores$Predictive)
summary(treatment.scores$Predictive)
levels(treatment.scores$Predictive)[levels(treatment.scores$Predictive)=="0"] <- "No"
levels(treatment.scores$Predictive)[levels(treatment.scores$Predictive)=="1"] <- "Yes"

#Yes
treatment.scores$Predictive <- as.character(treatment.scores$Predictive)
grp.predictyes <- treatment.scores[treatment.scores$Predictive == "Yes", ][chull(treatment.scores[treatment.scores$Predictive == "Yes", c("NMDS1", "NMDS2")]), ] 
#No
grp.predictno <- treatment.scores[treatment.scores$Predictive == "No", ][chull(treatment.scores[treatment.scores$Predictive == "No", c("NMDS1", "NMDS2")]), ]

#combine the hull data
hull.data.predict <- rbind(grp.predictyes, grp.predictno)  
str(hull.data.predict)

```

```{r Convex hulls - Type of Study}
# Type of study
unique(treatment.scores$TOS)

#Create convex hulls for the space occupied by each Type of study
#Hull loop
tos = as.character(unique(treatment.scores$TOS))
for(i in 1:length(tos)) {
  temp = tos[i]
  df = treatment.scores[treatment.scores$TOS == temp, ][chull(treatment.scores[treatment.scores$TOS == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.tos <- rbind(grp.Observational, grp.Experiment, grp.Metanalysis, grp.Theoretical, grp.Review)  

```

```{r Convex hulls - TT}
# Type of study
unique(treatment.scores$TT)

#Create convex hulls for the space occupied by each Type of study
#Hull loop
tt = as.character(unique(treatment.scores$TT))
for(i in 1:length(tt)) {
  temp = tt[i]
  df = treatment.scores[treatment.scores$TT == temp, ][chull(treatment.scores[treatment.scores$TT == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.tt <- rbind(grp.Morphological, grp.Behavioural, `grp.Life History`, grp.Physiological)  

```

```{r Convex hulls - Ecosystem}
### Ecosystem type
#Create convex hulls for the space occupied by each Ecosystem type
unique(treatment.scores$Ecosystem) #"Terrestrial" "Freshwater"  "Marine"      "Broad"

#Terrestrial
grp.t <- treatment.scores[treatment.scores$Ecosystem == "Terrestrial", ][chull(treatment.scores[treatment.scores$Ecosystem == "Terrestrial", c("NMDS1", "NMDS2")]), ] 
#Freshwater
grp.f <- treatment.scores[treatment.scores$Ecosystem == "Freshwater", ][chull(treatment.scores[treatment.scores$Ecosystem == "Freshwater", c("NMDS1", "NMDS2")]), ] 
#Marine
grp.m <- treatment.scores[treatment.scores$Ecosystem == "Marine", ][chull(treatment.scores[treatment.scores$Ecosystem == "Marine", c("NMDS1", "NMDS2")]), ]
#Broad
grp.b <- treatment.scores[treatment.scores$Ecosystem == "Multiple", ][chull(treatment.scores[treatment.scores$Ecosystem == "Multiple", c("NMDS1", "NMDS2")]), ]

#combine the hull data
hull.data.ecos <- rbind(grp.t, grp.f, grp.m, grp.b)  
str(hull.data.ecos)

```

```{r Convex hulls - Taxonomic Group}
# Taxonomic groups (broad)
#Create convex hulls for the space occupied by each Taxonomic group values
unique(treatment.scores$`Taxonomic Group`) 

#Taxonomic group - hull loop
taxgrp = as.character(unique(treatment.scores$`Taxonomic Group`))
for(i in 1:length(taxgrp)) {
  temp = taxgrp[i]
  df = treatment.scores[treatment.scores$`Taxonomic Group` == temp, ][chull(treatment.scores[treatment.scores$`Taxonomic Group` == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.taxgrp <- rbind(grp.Vertebrate, grp.Invertebrate, grp.Multiple, grp.Other, grp.Plants)  
str(hull.data.taxgrp)

```

```{r Convex hulls - Taxonomic specific}
# Taxonomic groups (specific)
#Create convex hulls for the space occupied by each Taxonomic values
unique(treatment.scores$Taxonomic) 

#Change some of the values
treatment.scores$Taxonomic <- as.factor(treatment.scores$Taxonomic)
levels(treatment.scores$Taxonomic)[levels(treatment.scores$Taxonomic)=="Broad"] <- "Multiple"
levels(treatment.scores$Taxonomic)[levels(treatment.scores$Taxonomic)=="Herps"] <- "Reptiles"
#"Birds"    "Insects"  "Mammals"  "Broad"    "Multiple" "Other"    "Fish"     "Herps"    "Plants"   "Plankton"
summary(treatment.scores$Taxonomic)

#Taxonomic group - hull loop
tax = as.character(unique(treatment.scores$Taxonomic))
for(i in 1:length(tax)) {
  temp = tax[i]
  df = treatment.scores[treatment.scores$Taxonomic == temp, ][chull(treatment.scores[treatment.scores$Taxonomic == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.taxon <- rbind(grp.Birds, grp.Insects, grp.Mammals, grp.Multiple, grp.Other, grp.Fish, grp.Reptiles, grp.Plants, grp.Plankton)  
head(hull.data.taxon)
```


## nMDS Plots

### A/ Global change driver

```{r nMDS Global Change Driver, echo=TRUE}
#nMDS plot for the assessmenet of global change (yes/no) and the drivers.

trait_nMDS_clim <- ggplot() + 
  geom_polygon(data=hull.data.gcd, 
               aes(x=NMDS1,y=NMDS2, fill= fct_relevel(`GlobalChange`, "Not assessed", "Habitat loss", "Climate change", "Multiple", "Invasion", "Exploitation"), 
                   group=fct_relevel(`GlobalChange`, "Not assessed", "Habitat loss", "Climate change", "Multiple", "Invasion", "Exploitation")), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores, aes(x=NMDS1,y=NMDS2, colour = fct_relevel(`GlobalChange`, "Not assessed", "Habitat loss", "Climate change", "Multiple", "Invasion", "Exploitation")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18), 
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  #guides(fill = guide_legend(order = 1), colour = guide_legend(order = 2)) +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Driver of change") + #name = "Global Change Assessed"
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Driver of change") #name = "Global Change Assessed"

trait_nMDS_clim

ggsave("trait_nMDS_clim.jpeg", plot = trait_nMDS_clim, width = 8, height = 8, dpi = 300)

```

### B/ Filter

```{r nMDS Filter, echo=FALSE}
#nMDS plot for the level of ecological filtering for which traits were applied

trait_nMDS_filt <- ggplot() + 
  geom_polygon(data=hull.data.filt,aes(x=NMDS1, y=NMDS2, fill= fct_relevel(filter, "Abiotic", "Biotic", "Dispersal", "Trophic"), group=fct_relevel(filter, "Abiotic", "Biotic", "Dispersal", "Trophic")),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_relevel(filter, "Abiotic", "Biotic", "Dispersal", "Trophic")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(plot.title=element_text(face="bold", size=18), # hjust = 0.05, vjust = -1 # margin = margin(t = 5, b = -25)
        axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Filter") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Filter")

trait_nMDS_filt

ggsave("trait_nMDS_filt.jpeg", plot = trait_nMDS_filt, width = 8, height = 8, dpi = 300)

```

### C/ Predictive studies

```{r nMDS C/ Predictive, echo=TRUE}
# nMDS for studies using traits for predictive/forecasting purposes

trait_nMDS_predict <- ggplot() + 
  geom_polygon(data=hull.data.predict, aes(x=NMDS1,y=NMDS2, fill=fct_relevel(Predictive, "Yes", "No"), group=fct_relevel(Predictive, "Yes", "No")), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores, aes(x=NMDS1,y=NMDS2, colour = fct_relevel(Predictive, "Yes", "No")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.8, name = "Predictive") + #name = "Predictive"
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.8, name = "Predictive") #name = "Predictive"

trait_nMDS_predict

ggsave("trait_nMDS_predict.jpeg", plot = trait_nMDS_predict, width = 8, height = 8, dpi = 300)


```

### D/ Study type

```{r nMDS D/ Study Type, echo=FALSE}
#nMDS plot for the type of study

trait_nMDS_tos <- ggplot() + 
  geom_polygon(data=hull.data.tos,aes(x=NMDS1,y=NMDS2,fill=fct_relevel(TOS, "Observational", "Experiment", "Metanalysis", "Review", "Theoretical"),group=fct_relevel(TOS, "Observational", "Experiment", "Metanalysis", "Review", "Theoretical")),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_relevel(TOS, "Observational", "Experiment", "Metanalysis", "Review", "Theoretical")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Study type") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Study type")

trait_nMDS_tos

ggsave("trait_nMDS_tos.jpeg", plot = trait_nMDS_tos, width = 8, height = 8, dpi = 300)

```


### E/ Ecosystem

```{r nMDS E/ Ecosystem, echo=TRUE}

#nMDS plot for the ecosystem type studied

trait_nMDS_ecos <- ggplot() + 
  geom_polygon(data=hull.data.ecos,aes(x=NMDS1,y=NMDS2, fill=fct_rev(`Ecosystem`), group=fct_rev(`Ecosystem`)), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_rev(`Ecosystem`)), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
   #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Ecosystem") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Ecosystem")
  
trait_nMDS_ecos

ggsave("trait_nMDS_ecos.jpeg", plot = trait_nMDS_ecos, width = 8, height = 8, dpi = 300)


```

### F/ Taxonomic Group

```{r nMDS F/ Taxonomic Group, echo=TRUE}
#nMDS plot for the broad taxonomic group studied

trait_nMDS_taxgrp <- ggplot() + 
  geom_polygon(data=hull.data.taxgrp,aes(x=NMDS1,y=NMDS2, fill= fct_relevel(`Taxonomic Group`, "Vertebrate", "Invertebrate", "Plants", "Other", "Multiple"), group= fct_relevel(`Taxonomic Group`, "Vertebrate", "Invertebrate", "Plants", "Other", "Multiple")),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_relevel(`Taxonomic Group`, "Vertebrate", "Invertebrate", "Plants", "Other", "Multiple")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Taxonomic group") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Taxonomic group")

trait_nMDS_taxgrp

ggsave("trait_nMDS_taxgrp.jpeg", plot = trait_nMDS_taxgrp, width = 8, height = 8, dpi = 300)


```

### G/ Trait type

```{r nMDS Taxonomic specific, echo=TRUE}
#nMDS plot for the specific taxonomic group studied

trait_nMDS_tt <- ggplot() + 
  geom_polygon(data=hull.data.tt,aes(x=NMDS1,y=NMDS2, fill= `TT`, group= `TT`),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = `TT`), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.7, name="Trait type") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.7, name="Trait type")

trait_nMDS_tt

ggsave("trait_nMDS_taxonsp.jpeg", plot = trait_nMDS_taxon, width = 8, height = 8, dpi = 300)

```


## Merge plots

Useful resources for aligning the plots, from the Wilke Lab:
https://wilkelab.org/cowplot/articles/aligning_plots.html 


```{r Merge nMDS plots, include=F, eval=F, echo=F}

#All 6 panels, grey background
review_nMDS_6 = plot_grid(trait_nMDS_clim, trait_nMDS_filt, trait_nMDS_predict, trait_nMDS_tos, trait_nMDS_ecos, trait_nMDS_taxgrp, labels = c("A", "B", "C", "D", "E", "F"), label_x = 0.12, vjust = 3.5, ncol = 2, rel_widths = c(1, 1), align = "v") #hjust = -7.5, label_x = c(0.12, 0.12, 0.12, 0.12, 0.12, 0.12)
review_nMDS_6
ggsave("review_nMDS_primary6.jpeg", plot = review_nMDS_6, width = 12, height = 10, dpi = 300)

#Alternative panel arrangements
#All 4 panels, grey background
#review_nMDS_4 = plot_grid(trait_nMDS_clim, trait_nMDS_filt, trait_nMDS_tos, trait_nMDS_ecos, labels = c("A", "B", "C", "D"), label_x = 0.10, vjust = 3.1, ncol = 2, rel_widths = c(1, 1),  align = "v")
#review_nMDS_4
#ggsave("review_nMDS_4whitelab.jpeg", plot = review_nMDS_4, width = 12, height = 7, dpi = 300)

```