---
title: "trait_review_nmds"
author: "Natasha Hardy"
date: "March 10, 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Workspace Install, include=FALSE}

getwd()
#install.packages("xlsx")
#library(devtools)
#install_github("vqv/ggbiplot")
#install.packages("goeveg")

```

```{r Workspace Load, include=FALSE}

#Data manip
library(tidyverse)
library(plyr)

#Load after
library(dplyr)
library(openxlsx)

#Stats / ordination
library(permute)
library(lattice)
library(vegan)
library(mvabund)

#Graphs
library(ggplot2)
library(scales)
library(grid)
library(gridExtra)
library(viridis)
library(PNWColors)
library(cowplot)
library(forcats)
library(goeveg)
library(ggrepel)

```


## UPLOAD CLEAN DATA FROM REPO

UPLOAD CLEANED DATA HERE & USE THIS DATA

```{r Initial Dataset}

#Upload selected dataset

#review_traits_use <- read_csv("review_traits4.0V.csv")
review_traits_use = read_csv("C:/Users/Helder/Documents/GitHub/traits_review_proj/Data/review_traits4.0V.csv") %>%
  select(DOI:ncol(.))
review_sites_use <- as.data.frame(review_traits_use[,1:10])
review_species_use <- review_traits_use[,11:ncol(review_traits_use)] #1411 trait columns

head(review_traits_use)
```


## Ordinations

Using Jaccard dissimilarity for binary data.

Note that the stress decreases as we decrease the number of observations, however, I'm concerned that it becomes so low as to indicate that we do not have enough observations vs. traits (data are too wide because #traits > # obs)

```{r Ordination USE, echo=F, eval=F}
#Don't run, run time = 3h

#Ordination based on using no replicates for the study
traits_ord_use <- metaMDS(review_species_use,
                          distance = "jaccard",
                          trymax = 1000,
                          maxits = 1000,
                          k = 4) #increase trymax = 50

##Model convergence issues
#Based on ordination goodness fit/stress checks, we could use k = 2-3.
#Don't want to redo this right now as it will take ~2h to rerun.
#Likely the data are zero inflated

plot(traits_ord_use) #plots species (here traits) as black dots, sites (here studies) as red crosses, not labelled

#Plot sites (row numbers) and species (traits) as labels #run the whole chunk of text below
ordiplot(traits_ord_use, type = "none") #creates blank plot
orditorp(traits_ord_use, display = "species", col = "black", air=0.01) #labels traits
orditorp(traits_ord_use, display = "sites", col = "red", air=0.01) #labels traits


ggsave("review_nmdsinitial.jpeg", width = 8, height = 8, dpi = 300)


saveRDS(traits_ord_use, "traits_ord_use.rds")

```



### UPLOAD ORD FILE

--> You can start directly here for further analysis pipeline

```{r UPLOAD ORDINATION RDS FILE}

#Initial dummy data
traits_ord_use <- readRDS("C:/Users/Helder/Documents/GitHub/traits_review_proj/Output-Figs/traits_ord_use.rds")
View(traits_ord_use)
#Old model did not converge

#Secondary trait classification filled in
#traits_ord_use <- readRDS("traits_ord_2use.rds")
#primary trait classification filled in
traits_ord <- readRDS("traits_ord_1use.rds")

```

## Checking ordination/nMDS output

This is a helpful tutorial on checking nMDS goodness of fit, model convergence, stress and dimensionality: https://ourcodingclub.github.io/tutorials/ordination/ 
Notes:
For the initial dummy dataset - stress is extremely low, likely there are insufficient data (n = 727 studies/rows) for the number of traits (n = 1411) included in this analysis.
For the primary traits dataset - the inflection point on the stress vs dimensions plot is between 2-3, so I am selecting k= 3 

```{r Ordination - checks}
#Run this for each ordination
#Check fit
gof <- goodness(traits_ord_use)

#Can visualise goodness of fit
stressplot(traits_ord_use, pch = 19, cex=0.75, l.col = "tomato", p.col = "skyblue")

#Now, we will be computing the stress for the dimensions 1, 2, 3, 4, 5, …. and explain how many dimensions are required to obtain a good fit.
#REMEMBER TO ADJUST THE INITIAL SP MATRIX
stress_vec <- numeric(10)
for(i in seq(10)){
  stress_vec[i] <- metaMDS(review_species_use,
                          distance = "jaccard",
                          k=i)$stress
}

#Plot the stress vs. number of dimensions needed
plot(seq(10),stress_vec, type = 'o', ylab = "Stress", xlab = "Number of dimensions",
     col="tomato", pch=19)




#Plot goodness of fit of points
#plot(traits_ord, display = "sites", type = "n")
#points(traits_ord)

```

```{r nMDS stress check, eval=FALSE}
#Don't run, too time intensive.
#We know from output above that the model is underdispersed and we have fit too few rows vs. variables.

#dimcheckMDS(review_species_use,
#            distance = "jaccard",
#            trymax = 1000,
#            k = 3)

```


## GO TO GRAPHICS PIPELINE

For calculating data and treatment scores, and convex hulls around datapoints.

### Extract NMDS coordinates and associate with co-variates/grouping factors

```{r Data scores for plotting - USE}

#Extract NMDS coordinates and associate with co-variates/grouping factors

#Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores <- as.data.frame(scores(traits_ord_use))  #, "species"

#create a column of site names, from the rownames of data.scores
data.scores$points <- rownames(data.scores)

#bind treatment labels and score values
treatment.scores <- cbind(review_sites_use, data.scores)

#Check
str(treatment.scores)  
#Awesome

#ALSO
#OR for species scores
species.scores <- as.data.frame(scores(traits_ord_use, "species"))
#create a column of species names:
species.scores$species <- rownames(species.scores)
#str(data.scores)

```

### Convex hull calculations
For each ordination and set of grouping variables input to data scores chunk

```{r Convex hulls - Global Change}
#Global Change yes/no
#Create convex hulls for the space occupied by Global Change yes/no
unique(treatment.scores$`Global Change`) #no/yes
treatment.scores$`Global Change` <- as.factor(treatment.scores$`Global Change`)
summary(treatment.scores$`Global Change`)
levels(treatment.scores$`Global Change`)[levels(treatment.scores$`Global Change`)=="no"] <- "No"
levels(treatment.scores$`Global Change`)[levels(treatment.scores$`Global Change`)=="yes"] <- "Yes"

#Yes
grp.climyes <- treatment.scores[treatment.scores$`Global Change` == "Yes", ][chull(treatment.scores[treatment.scores$`Global Change` == "yes", c("NMDS1", "NMDS2")]), ] 
#No
grp.climno <- treatment.scores[treatment.scores$`Global Change` == "No", ][chull(treatment.scores[treatment.scores$`Global Change` == "no", c("NMDS1", "NMDS2")]), ] 

#combine the hull data
hull.data.clim <- rbind(grp.climyes, grp.climno)  
str(hull.data.clim)
```

```{r Convex hulls - Global Change Driver}
#Global Change Drivers
#Create convex hulls for the space occupied by each Global Change Driver
unique(treatment.scores$`Global Change Driver`)

#Change driver labels
treatment.scores$`Global Change Driver` <- as.factor(treatment.scores$`Global Change Driver`)
summary(treatment.scores$`Global Change Driver`)

treatment.scores$`Global Change Driver` <- as.factor(treatment.scores$`Global Change Driver`)
levels(treatment.scores$`Global Change Driver`)[levels(treatment.scores$`Global Change Driver`)==0] <- "Not assessed"
levels(treatment.scores$`Global Change Driver`)[levels(treatment.scores$`Global Change Driver`)=="Global Change Broad"] <- "Multiple"
levels(treatment.scores$`Global Change Driver`)[levels(treatment.scores$`Global Change Driver`)=="Global Change Multiple"] <- "Multiple"
levels(treatment.scores$`Global Change Driver`)[levels(treatment.scores$`Global Change Driver`)=="Habitat Degredation"] <- "Habitat loss"
levels(treatment.scores$`Global Change Driver`)[levels(treatment.scores$`Global Change Driver`)=="Climate Change"] <- "Climate change"

#Global Change Drivers - hull loop
gcd = as.character(unique(treatment.scores$`Global Change Driver`))
for(i in 1:length(gcd)) {
  temp = gcd[i]
  df = treatment.scores[treatment.scores$`Global Change Driver` == temp, ][chull(treatment.scores[treatment.scores$`Global Change Driver` == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.gcd <- rbind(`grp.Not assessed`,`grp.Climate change`, grp.Exploitation, grp.Multiple, `grp.Habitat loss`, grp.Invasion)  
str(hull.data.gcd)

```

```{r Convex hulls - Filter}
# Level of filtering

#Make following value changes
#"Fundamental" --> Abiotic
#"Physical" --> Dispersal 
#"Ecological"  --> Biotic
#"Trophic" = "Trophic"
treatment.scores$Filter <- as.factor(treatment.scores$Filter)
levels(treatment.scores$Filter)[levels(treatment.scores$Filter)=="Fundamental"] <- "Abiotic"
levels(treatment.scores$Filter)[levels(treatment.scores$Filter)=="Physical"] <- "Dispersal"
levels(treatment.scores$Filter)[levels(treatment.scores$Filter)=="Ecological"] <- "Biotic"

#Create convex hulls for the space occupied by each Type of study
#Hull loop
filt = as.character(unique(treatment.scores$Filter))
for(i in 1:length(filt)) {
  temp = filt[i]
  df = treatment.scores[treatment.scores$Filter == temp, ][chull(treatment.scores[treatment.scores$Filter == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.filt <- rbind(grp.Abiotic, grp.Biotic, grp.Trophic, grp.Dispersal)  
str(hull.data.filt)

#Check data distribution
#unique(hull.data.filt$Filter)
#summary(treatment.scores$Filter)

```

```{r Convex hulls - Predictive}
#Predictive/Forecasting studies
#Create convex hulls for the space occupied by Predictive yes/no
unique(treatment.scores$Predictive) #no/yes

#Modification of any values
treatment.scores$Predictive <- as.factor(treatment.scores$Predictive)
summary(treatment.scores$Predictive)
levels(treatment.scores$Predictive)[levels(treatment.scores$Predictive)=="0"] <- "No"
levels(treatment.scores$Predictive)[levels(treatment.scores$Predictive)=="1"] <- "Yes"

#Yes
treatment.scores$Predictive <- as.character(treatment.scores$Predictive)
grp.predictyes <- treatment.scores[treatment.scores$Predictive == "Yes", ][chull(treatment.scores[treatment.scores$Predictive == "Yes", c("NMDS1", "NMDS2")]), ] 
#No
grp.predictno <- treatment.scores[treatment.scores$Predictive == "No", ][chull(treatment.scores[treatment.scores$Predictive == "No", c("NMDS1", "NMDS2")]), ]

#combine the hull data
hull.data.predict <- rbind(grp.predictyes, grp.predictno)  
str(hull.data.predict)

```

```{r Convex hulls - Type of Study}
# Type of study
unique(treatment.scores$TOS)
#TModel --> Theory
treatment.scores$TOS <- as.factor(treatment.scores$TOS)
levels(treatment.scores$TOS)[levels(treatment.scores$TOS)=="TModel"] <- "Theory"
#Create convex hulls for the space occupied by each Type of study
#Hull loop
tos = as.character(unique(treatment.scores$TOS))
for(i in 1:length(tos)) {
  temp = tos[i]
  df = treatment.scores[treatment.scores$TOS == temp, ][chull(treatment.scores[treatment.scores$TOS == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.tos <- rbind(grp.Observational, grp.Experiment, grp.Metanalysis, grp.Theory, grp.Review)  
```

```{r Convex hulls - Ecosystem}
### Ecosystem type
#Create convex hulls for the space occupied by each Ecosystem type
unique(treatment.scores$Ecosystem) #"Terrestrial" "Freshwater"  "Marine"      "Broad"

#Terrestrial
grp.t <- treatment.scores[treatment.scores$Ecosystem == "Terrestrial", ][chull(treatment.scores[treatment.scores$Ecosystem == "Terrestrial", c("NMDS1", "NMDS2")]), ] 
#Freshwater
grp.f <- treatment.scores[treatment.scores$Ecosystem == "Freshwater", ][chull(treatment.scores[treatment.scores$Ecosystem == "Freshwater", c("NMDS1", "NMDS2")]), ] 
#Marine
grp.m <- treatment.scores[treatment.scores$Ecosystem == "Marine", ][chull(treatment.scores[treatment.scores$Ecosystem == "Marine", c("NMDS1", "NMDS2")]), ]
#Broad
grp.b <- treatment.scores[treatment.scores$Ecosystem == "Broad", ][chull(treatment.scores[treatment.scores$Ecosystem == "Broad", c("NMDS1", "NMDS2")]), ]

#combine the hull data
hull.data.ecos <- rbind(grp.t, grp.f, grp.m, grp.b)  
str(hull.data.ecos)

```

```{r Convex hulls - Taxonomic Group}
# Taxonomic groups (broad)
#Create convex hulls for the space occupied by each Taxonomic group values
unique(treatment.scores$`Taxonomic Group`) 

#Taxonomic group - hull loop
taxgrp = as.character(unique(treatment.scores$`Taxonomic Group`))
for(i in 1:length(taxgrp)) {
  temp = taxgrp[i]
  df = treatment.scores[treatment.scores$`Taxonomic Group` == temp, ][chull(treatment.scores[treatment.scores$`Taxonomic Group` == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.taxgrp <- rbind(grp.Vertebrate, grp.Invertebrate, grp.Multiple, grp.Other, grp.Plants)  
str(hull.data.taxgrp)

```

```{r Convex hulls - Taxonomic specific}
# Taxonomic groups (specific)
#Create convex hulls for the space occupied by each Taxonomic values
unique(treatment.scores$Taxonomic) 

#Change some of the values
treatment.scores$Taxonomic <- as.factor(treatment.scores$Taxonomic)
levels(treatment.scores$Taxonomic)[levels(treatment.scores$Taxonomic)=="Broad"] <- "Multiple"
levels(treatment.scores$Taxonomic)[levels(treatment.scores$Taxonomic)=="Herps"] <- "Reptiles"
#"Birds"    "Insects"  "Mammals"  "Broad"    "Multiple" "Other"    "Fish"     "Herps"    "Plants"   "Plankton"
summary(treatment.scores$Taxonomic)

#Taxonomic group - hull loop
tax = as.character(unique(treatment.scores$Taxonomic))
for(i in 1:length(tax)) {
  temp = tax[i]
  df = treatment.scores[treatment.scores$Taxonomic == temp, ][chull(treatment.scores[treatment.scores$Taxonomic == temp, c("NMDS1", "NMDS2")]), ]
  assign(paste0('grp.',temp), df)
}

#combine the hull data
hull.data.taxon <- rbind(grp.Birds, grp.Insects, grp.Mammals, grp.Multiple, grp.Other, grp.Fish, grp.Reptiles, grp.Plants, grp.Plankton)  
head(hull.data.taxon)
```


## nMDS Plots

### A/ Global change driver

```{r nMDS Global Change Driver, echo=TRUE}
#nMDS plot for the assessmenet of global change (yes/no) and the drivers.

trait_nMDS_clim <- ggplot() + 
  geom_polygon(data=hull.data.gcd, 
               aes(x=NMDS1,y=NMDS2, fill= fct_relevel(`Global Change Driver`, "Not assessed", "Habitat loss", "Climate change", "Multiple", "Invasion", "Exploitation"), 
                   group=fct_relevel(`Global Change Driver`, "Not assessed", "Habitat loss", "Climate change", "Multiple", "Invasion", "Exploitation")), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores, aes(x=NMDS1,y=NMDS2, colour = fct_relevel(`Global Change Driver`, "Not assessed", "Habitat loss", "Climate change", "Multiple", "Invasion", "Exploitation")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18), 
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  #guides(fill = guide_legend(order = 1), colour = guide_legend(order = 2)) +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Driver of change") + #name = "Global Change Assessed"
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Driver of change") #+ #name = "Global Change Assessed"
  #geom_text_repel(data = species.scores,
  #                aes(x = NMDS1, y = NMDS2, label = species#, alpha = 0.8, size = 5
  #                    ))

trait_nMDS_clim

ggsave("trait_nMDS_clim.jpeg", plot = trait_nMDS_clim, width = 8, height = 8, dpi = 300)

```

### B/ Filter

```{r nMDS Filter, echo=FALSE}
#nMDS plot for the level of ecological filtering for which traits were applied

trait_nMDS_filt <- ggplot() + 
  geom_polygon(data=hull.data.filt,aes(x=NMDS1, y=NMDS2, fill= fct_relevel(Filter, "Abiotic", "Biotic", "Dispersal", "Trophic"), group=fct_relevel(Filter, "Abiotic", "Biotic", "Dispersal", "Trophic")),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_relevel(Filter, "Abiotic", "Biotic", "Dispersal", "Trophic")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(plot.title=element_text(face="bold", size=18), # hjust = 0.05, vjust = -1 # margin = margin(t = 5, b = -25)
        axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Filter") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Filter")

trait_nMDS_filt

ggsave("trait_nMDS_filt.jpeg", plot = trait_nMDS_filt, width = 8, height = 8, dpi = 300)

```

### C/ Predictive studies

```{r nMDS C/ Predictive, echo=TRUE}
# nMDS for studies using traits for predictive/forecasting purposes

trait_nMDS_predict <- ggplot() + 
  geom_polygon(data=hull.data.predict, aes(x=NMDS1,y=NMDS2, fill=fct_relevel(Predictive, "Yes", "No"), group=fct_relevel(Predictive, "Yes", "No")), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores, aes(x=NMDS1,y=NMDS2, colour = fct_relevel(Predictive, "Yes", "No")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.8, name = "Predictive") + #name = "Predictive"
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.8, name = "Predictive") #name = "Predictive"

trait_nMDS_predict

ggsave("trait_nMDS_predict.jpeg", plot = trait_nMDS_predict, width = 8, height = 8, dpi = 300)


```

### D/ Study type

```{r nMDS D/ Study Type, echo=FALSE}
#nMDS plot for the type of study

trait_nMDS_tos <- ggplot() + 
  geom_polygon(data=hull.data.tos,aes(x=NMDS1,y=NMDS2,fill=fct_relevel(TOS, "Observational", "Experiment", "Metanalysis", "Review", "Theory"),group=fct_relevel(TOS, "Observational", "Experiment", "Metanalysis", "Review", "Theory")),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_relevel(TOS, "Observational", "Experiment", "Metanalysis", "Review", "Theory")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Study type") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name = "Study type")

trait_nMDS_tos

ggsave("trait_nMDS_tos.jpeg", plot = trait_nMDS_tos, width = 8, height = 8, dpi = 300)

```


### E/ Ecosystem

```{r nMDS E/ Ecosystem, echo=TRUE}

#nMDS plot for the ecosystem type studied

trait_nMDS_ecos <- ggplot() + 
  geom_polygon(data=hull.data.ecos,aes(x=NMDS1,y=NMDS2, fill=fct_rev(`Ecosystem`), group=fct_rev(`Ecosystem`)), alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_rev(`Ecosystem`)), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
   #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Ecosystem") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Ecosystem")
  
trait_nMDS_ecos

ggsave("trait_nMDS_ecos.jpeg", plot = trait_nMDS_ecos, width = 8, height = 8, dpi = 300)


```

### F/ Taxonomic Group

```{r nMDS F/ Taxonomic Group, echo=TRUE}
#nMDS plot for the broad taxonomic group studied

trait_nMDS_taxgrp <- ggplot() + 
  geom_polygon(data=hull.data.taxgrp,aes(x=NMDS1,y=NMDS2, fill= fct_relevel(`Taxonomic Group`, "Vertebrate", "Invertebrate", "Plants", "Other", "Multiple"), group= fct_relevel(`Taxonomic Group`, "Vertebrate", "Invertebrate", "Plants", "Other", "Multiple")),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = fct_relevel(`Taxonomic Group`, "Vertebrate", "Invertebrate", "Plants", "Other", "Multiple")), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.justification = c(0,0.5),
        #panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  #guides(fill = FALSE, colour = FALSE) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Taxonomic group") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.8, end = 0.2, name="Taxonomic group")

trait_nMDS_taxgrp

ggsave("trait_nMDS_taxgrp.jpeg", plot = trait_nMDS_taxgrp, width = 8, height = 8, dpi = 300)


```

### EXTRA Taxonomic (Specific)

```{r nMDS Taxonomic specific, echo=TRUE}
#nMDS plot for the specific taxonomic group studied

trait_nMDS_taxon <- ggplot() + 
  geom_polygon(data=hull.data.taxon,aes(x=NMDS1,y=NMDS2, fill= `Taxonomic`, group= `Taxonomic`),alpha=0.30) + # add the convex hulls
  geom_point(data=treatment.scores,aes(x=NMDS1,y=NMDS2, colour = `Taxonomic`), size=2) + # add the point markers
  coord_equal() +
  theme_bw()  +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        panel.background = element_rect(fill = "lightgrey"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) +
  scale_colour_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.7, name="Taxonomic") +
  scale_fill_viridis(option="magma", discrete = TRUE, begin = 0.2, end = 0.7, name="Taxonomic")

trait_nMDS_taxon

ggsave("trait_nMDS_taxonsp.jpeg", plot = trait_nMDS_taxon, width = 8, height = 8, dpi = 300)

```


## Merge plots

Useful resources for aligning the plots, from the Wilke Lab:
https://wilkelab.org/cowplot/articles/aligning_plots.html 


```{r Merge nMDS plots, include=F, eval=F, echo=F}

#All 6 panels, grey background
review_nMDS_6 = plot_grid(trait_nMDS_clim, trait_nMDS_filt, trait_nMDS_predict, trait_nMDS_tos, trait_nMDS_ecos, trait_nMDS_taxgrp, labels = c("A", "B", "C", "D", "E", "F"), label_x = 0.12, vjust = 3.5, ncol = 2, rel_widths = c(1, 1), align = "v") #hjust = -7.5, label_x = c(0.12, 0.12, 0.12, 0.12, 0.12, 0.12)
review_nMDS_6
ggsave("review_nMDS_6whitelab.jpeg", plot = review_nMDS_6, width = 12, height = 10, dpi = 300)


#All 4 panels, grey background
review_nMDS_4 = plot_grid(trait_nMDS_clim, trait_nMDS_filt, trait_nMDS_tos, trait_nMDS_ecos, labels = c("A", "B", "C", "D"), label_x = 0.10, vjust = 3.1, ncol = 2, rel_widths = c(1, 1),  align = "v")
review_nMDS_4
ggsave("review_nMDS_4whitelab.jpeg", plot = review_nMDS_4, width = 12, height = 7, dpi = 300)

```

#Useful for troubleshoot legend alignment
https://github.com/wilkelab/cowplot/issues/33

## Simper Analyses
##Similarity Percentage Analysis

Understanding the output in a nutshell:
https://rdrr.io/rforge/vegan/man/simper.html

Notes from rdocumentation: https://www.rdocumentation.org/packages/vegan/versions/2.4-2/topics/simper

The simper functions performs pairwise comparisons of groups of sampling units and finds the average contributions of each species to the average overall Bray-Curtis dissimilarity.

The function displays most important species for each pair of groups. These species contribute at least to 70 % of the differences between groups. The function returns much more extensive results which can be accessed directly from the result object (see section Value). Function summary transforms the result to a list of data frames. With argument ordered = TRUE the data frames also include the cumulative contributions and are ordered by species contribution.

The results of simper can be very difficult to interpret. The method very badly confounds the mean between group differences and within group variation, and seems to single out variable species instead of distinctive species (Warton et al. 2012). Even if you make groups that are copies of each other, the method will single out species with high contribution, but these are not contributions to non-existing between-group differences but to within-group variation in species abundance.

Most of the code for saved SIMPER files are currently on PACIFIC computer.

```{r SIMPER Ecosystem, eval=F, echo=T}

simper_ecos <- simper(review_species_use, review_sites_use$Ecosystem, permutations = 1000)

summary(simper_ecos)

```

```{r SIMPER Filter, eval=F, echo=T}

simper_filter <- simper(review_species_use, review_sites_use$Filter, permutations = 1000)

summary_filter <- summary(simper_filter, ordered=TRUE)

#Tried and failed to create multi-sheet xlsx document out of SIMPER output
#str(summary_filter)
#str(simper_filter)

#write.csv(simper_filter$Fundamental_Ecological, "SIMPER_Filter_Fundamental_Ecological.csv")

#output <- 1
#for (i in length(summary_filter)){
#   write.xlsx(x=summary_filter[[i]],file="summary_filter.xlsx", sheetName=paste("sheet",output,sep=""),append=T)
#   output <- output + 1
#}

```

```{r SIMPER Global Change, eval=F, echo=T}

simper_gc <- simper(review_species_use, review_sites_use$`Global Change`, permutations = 1000)

summary(simper_gc)

```

```{r SIMPER Type of Study, eval=F, echo=T}

simper_tos <- simper(review_species_use, review_sites_use$TOS, permutations = 1000)

summary(simper_tos)

```

## Don't run code below here

```{r Gg.arrange plot merging, include=F, eval=F, echo=T}

#Potential gg.arrange work around
#gP <- ggplotGrob(trait_nMDS_clim)
#gQ <- ggplotGrob(trait_nMDS_ecos)
#review_grid = grid::grid.newpage(); grid.arrange(rbind(gQ, gP))
#ggexport(review_grid, filename = "review_grid.jpeg")



```


#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
#`eval = FALSE` ...